Fractal Modeling for Machine Learning

### 2012-11-05 fractal paper/src/fractal.tex
...

\documentclass[10pt]{article}

@@

\DeclareMathOperator*{\argmin}{arg\,min}

@@

\begin{abstract}
\noindent We present a new algorithm for fitting fractals to data. We use this algorithm to investigate the use of fractal models in a machine learning context. Fractals are mathematical objects that display great detail and complexity despite having simple descriptions. It has often been suggested that fractals model many natural and other real-world structures better than the shapes of classical geometry \cite{mandelbrot1982fractal}.
\end{abstract}

\section{Introduction}

Fractals are mathematical objects (generally sets or measures) with certain remarkable properties. What constitutes a fractal is not precisely defined\footnote{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}, but the following properties are common:

\begin{description}
  \item[Self similarity] A part is a scaled down copy of the whole. See for instance the Sierpinski triangle (fig !@). It consists of three triangular shapes which are scaled down copies of the whole (as such these three components each also consist of scaled down copies of themselves, and so on).
  \item[Infinitely small detail] 'Zooming in' reveals ever finer detail. In the case of the Sierpinski triangle we will simply see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
  \item[Non-integer dimension] for instance the Sierpinski triangle has dimension of about $1.58$. This is a counter-intuitive notion, that is not simple to explain. We refer the reader to \cite{schroeder2009fractals} for an intuitive explanation.
\end{description}

\begin{figure}[h!]
  \centering
  \begin{subfigure}[b]{\textwidth}@!
    %\includegraphics[width=\textwidth]{}
    \label{fig:sierpinski}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}@!
   % \includegraphics[width=\textwidth]{}
    \label{fig:sierpinski-measure}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}@!
    %\includegraphics[width=\textwidth]{}
    \label{fig:ifs-measure}
  \end{subfigure}
  \caption{(\ref{fig:sierpinski}) The Sierpinski gasket, a simple example fractal. Starting with any initial image, we can compose three transformations of the original. The iteration of this map converges to the Sierpinski gasket. The initial image does not influence the end result. (\ref{fig:sierpinski-measure}) Starting with probability measures rather then images, we can define a fractal probability measure. Again the choice of initial distribution doesn't matter. In this example we have assigned one of the transformations a higher weight than the other two. (\ref{fig:ifs-measure})If we change the transformations, the limit set changes too. Here we show some probability distributions that are described in this way. Note that simple shapes such as the triangle, the square and the line can also be described like this.}  
\end{figure}

The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and mostly considered pathological. It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena. Mandelbrot put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:

@@

Fractal geometry has been used in many fields, including physics @! \cite{}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}. The first connection between fractals and the natural world was in the form of tangible objects such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well. The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta have all been studied as fractals to great effect.

@@

This problem was investigated thoroughly in the context of image compression \cite{hart1996fractal}, but has received relatively little attention in the area of statistics, and fitting a fractal model to data rather than an image. In particular the case of data in arbitrary dimensions has been left largely uninvestigated.

\subsection{Previous research}
[@! Review the literature that does exist]

\section{Methods}

In this paper we focus on numerical data an arbitrary number of dimensions. Each instance in our dataset is represented by a fixed number of numerical features. Call our dataset $X$, our instance $x \in X$ and $x \in {\mathbb R}^d$. We will assume that our dataset represents iid samples from a probability distribution over ${\mathbb R}^d$ which we will attempt to model.

\subsection{Iterated Function Systems}

@@

We will use the following notation. Let $S_1$, $S_2$, \ldots, $S_n$ be a set of contractive transformations on ${\mathbb R}^d$. (We will later limit the family of allowable functions, but from a mathematical point of view we only require them to be contractive). There exists a single set $Q$ that satisfies the following property:
\begin{align}
Q = S_1(Q) \cup S_2(Q) \cup \ldots \cup S_k(Q) \label{scaling law}
\end{align}

@@

\subsection{Representing IFSs}

@@

The method described below can easily be extended to other families of transformations such as affine transformations or neural networks. The only requirement is that we must have some way of finding a transformation $F$ given two ordered point sets ${x_i}, {y_i} \subseteq {\mathbb R}^d$ so that $|F(x_i) - y_i|$ is minimized. For similarities there is a direct and optimal method. [@! cite]

\subsection{Drawing points from an IFS distribution}

@@

It should be mentioned that the sequence of points generated this way dos not represent a sequence of independent draws. Any two successive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$ for some $i$. This is not true for two independently drawn points. For our experiments we generate a sequence and then shuffle the data to remove this dependency where it is undesirable.

@@

\subsection{Code space}
\label{sec:code-space}
[@! picture]

@@

Since each $S_i$ is contractive, we can let the length of $\sigma$ go to infinity and get 
\[
S_\sigma(X) = x_\sigma  
\] 
That is, the sequence of functions maps each $X$ to a unique point $x_\sigma$. In the case of probability measures, every probability measure is mapped to a probability measure that gives the single point $x_\sigma$ full probability mass. Figure [@!] illustrates the concept, an Iterated Function System assigns a code to each point on its support. Finite codes map to subsets of the support. Since the size of the subsets decreases exponentially we can talk about the endpoint of a finite coee without losing much accuracy. We will take the point to which the origin is mapped by $S_\sigma$ as the canonical choice of the endpoint of finite $\sigma$.

Finding the endpoint for given code is straightforward, but we will also require the inverse process: given a point $x$ on or near the support of some IFS, we wish to find $\sigma$. To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions. The $\sigma$ which assigns $x$ the greatest probability is chosen.

Since the variance of the distribution drops off exponentially as the length $l$ is increased, this method will fail if points are too far off the support of the IFS (even if logarithmic probabilities are used). To increase the range of the method we fall back on the distance from the point to the mean of the distribution. If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to the point.

A word of caution on codes is appropriate: the ordering is often the reverse of what our intuition suggests. If we number the three components of the Sierpinski triangle clockwise from the top 0, 1 and 2, we might number a small section of the triangle by first moving into the top subtriangle (applying transformation 0), then into the left subtriangle (2) of that and then into the right subtriangle (1) of that. Intuitively, this is the most natural way of coding sections of the Sierpinski triangle. [@! fig]

@@

Finally, we wish to note that a line segment, the square, the cube and higher dimensional analogues can be represented as simple IFSs of $2^d$ components. For instance, for the unit line, we map the whole onto its lower and upper half with two components. The codes of this IFS are binary strings and the subsets they code are the diadic intervals. For the square, we can map the whole onto each one of its quadrants and get quaternary codes. Using this mechanism we can always map any dataset into the bi-unit hypercube and find a straightforward coding fot the instances in $2^d$-ary codes. What fractal fitting offers us is a potential for a subspace that fits the data more compactly and gives us a good code of lower arity.

\subsection{Evaluating models}

@@

Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added. Since the Sierpinski gasket has dimension less than two, the probability that the point will still fall within the Sierpinski gasket will be zero. Thus all points will have probability zero under the Sierpinski gasket and its likelihood will be zero. Intuitively, the Sierpinski gasket is a good model for this data.

For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff distance  $H(X, Y)$ between two sets $X$ and $Y$ is defined as follows:
\[ 
H'(X, Y) = \max_{x\in X} \min_{y\in Y} d(x, y) 
\]
\[
H(X, Y) = \max\left \{H(X, Y), H(Y, X)\right\}
\]
 (for infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$)

The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the second player will choose his starting point to maximize that distance.

Hausdorff distance is a common tool in dealing with the fractal inverse problem [@! cite]. We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model. Since the Hausdorff distance has not been substantiated as a statistical measure, we can only use it as an intermediate indication of model quality. To prove the utility of IFS models we will use them to solve three classical machine learning problems (classification, dimension reduction and clustering).

We estimate the uncertainty in the estimate of the Hausdorff distance by repeating the process $50$ times and calculating the sample standard deviation.

\subsection{Dimension}

The curse of dimensionality is a blanket term for many effects related the analysis of high dimensional data. The main cause of these problems is that as dimensionality increases, the volume of a subspace (such as the unit ball) increases exponentially with it. This means, for instance, that vast amounts of samples are required to maintain the same level of coverage in a high dimensional space as compared to a low dimensional one.

@@

\subsection{Analogous model: Mixture of Gaussians}

To study the IFS as a statistical model, it helps to have a commonly used analogous model to compare it with. We use the Mixture of Gaussians (MoG) model. This model consists of $n$ multivariate Gaussians over ${\mathbb R}^d$ each represented by a mean $\mu \in {\mathbb R}^D$ and a $d \times d$ covariance matrix $\Sigma$. Each component also has a prior $w_i$. The complete model describes the following probability distribution:
\[
p(X) = \sum_{i \in [1,n]} w_i N(x|\mu_i, \Sigma)
\]
Since every multivariate Gaussian can also be described as an Affine transformation of the standard normal distribution (mean at the origin and $\Sigma$ the identity matrix), we can also represent the mixture of Gaussians as $n$ affine transformations and $n$ weights, just as the IFS:
\[
p(X) = \sum_{i \in [1,n]} w_i S_i(N(\mu_i, \Sigma))(X)
\]
This gives us a classical model with a similar representation. Note that the representation would be equal if we used similarities for the MoG model. However since an MVN transformed with uniform scaling is rotationally symmetric that would make all the rotational parameters useless. For this reason we will allow the MoG model full affine transformations as parameters and the IFS model only similarities. Even with this discrepancy, we still consider the MoG to be a reasonable analog.

@@

\begin{itemize}
  \item Expectation: Assign each data point a probability distribution over the components of the MoG model proportional to the likelihood of the model given the datapoint.
  \item Maximization: Estimate the parameters for each component based on the probability weights calculated in the previous step.  
\end{itemize}

@@

\section{Fitting IFS models to data}

@@

\subsection{The basic algorithm}

@@

\subsection{Improvements}

\subsubsection{Centering and scaling the data}

@@

\subsubsection{Using all depths}

@@

\subsubsection{Multiple matches}

@@

\subsubsection{Weights}

@@

\subsubsection{Edge cases}

@@

\section{Results}

@@

\subsection{Datasets}

@@

\subsection{Model fitting and dimension scaling}

@@

\subsection{Classification}

\subsubsection{Bayesian}

\subsubsection{Single model}

An alternative way to construct a classifier with IUFS modelling is to learn a single model for the whole dataset, ignoring class labels. For some datasets, the set of all points may present a more coherent and simple distribution than the sets split by class label. Another advantage of this approach is that we only have to learn one model, which is cheaper both in training and in evaluation.

The model provides us with a mapping from the data space (${\mathbb R}^d$) to code space (${1, \ldots, k}^d$). To turn this model into a classifier we proceed as follows. We build a code tree such that each possible code represents a path from the root to the leaf node. We then traverse this tree for each point in the training set, maintaining a frequency count fora each class observed at each node. Thus, the root node gives us the frequency distribution over the classes for the whole dataset, and each node below it refines that to a distribution over a section of the dataset whose code starts with a given prefix. This turns each node into a very basic classifier, which simply returns the class with the highest frequency count. The root node will give us the classifier that uses nothing mode than class frequencies and each node below it will use more information about the neighbourhood of the point.

@@

\section{Conclusions}

>>>>

\documentclass[10pt,a4paper,oneside]{article}

@@

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{argmax}

@@

\begin{abstract}
\noindent We present an algorithm for fitting fractals to data and we investigate the use of fractal models in machine learning. Fractals are mathematical objects that display great detail and complexity despite having simple descriptions. It has been suggested that fractals model many natural and other real-world structures better than the shapes of classical geometry \cite{mandelbrot1982fractal}.
\end{abstract}

\begin{figure}[b]
  \centering
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=\textwidth]{../img/sierpinski.pdf}
    \caption{}
    \label{fig:sierpinski}
  \end{subfigure}  
  \hspace{0.03\textwidth}
  \begin{subfigure}[b]{0.3\textwidth}
   \includegraphics[width=\textwidth]{../img/sierpinski-evo.pdf}
    \caption{}
  \end{subfigure}
  \hspace{0.03\textwidth}
  \begin{subfigure}[b]{0.3\textwidth}
    \includegraphics[width=0.45\textwidth]{../img/sierpinski-off.png}
    \includegraphics[width=0.45\textwidth]{../img/random013.png} \\
    \includegraphics[width=0.45\textwidth]{../img/random010.png}
    \includegraphics[width=0.45\textwidth]{../img/mvn1k.png}
    \caption{}
  \end{subfigure}
  \caption{\small (a) The Sierpinski triangle, a simple example fractal. (b) Constructing \textit{(a)} from an arbitrary initial image. (c) Examples of IFS probability distributions.}
  \label{fig:examples}
\end{figure}

Fractals are mathematical objects with certain remarkable properties. What constitutes a fractal is not precisely defined\footnotemark, but the following properties are common:

\begin{description}
  \item[Self similarity] A part is a scaled down copy of the whole. See for instance the Sierpinski triangle (fig. \ref{fig:sierpinski}). It consists of three triangular shapes which are scaled down copies of itself.
  \item[Infinitely small detail] 'Zooming in' reveals ever finer detail. In the case of the Sierpinski triangle we will see the same shape recurring again and again, but fractals like the Mandelbrot set reveal a great variety of shapes.
  \item[Non-integer dimension] The Sierpinski triangle, for example, has a dimension of $1.58$. This is a counter-intuitive notion, that is not simple to explain. We refer the reader to \cite{schroeder2009fractals} for an intuitive explanation.
\end{description}

The earliest examples of fractals were described around the beginning of the twentieth century as counterexamples to conjectures in planar geometry and analysis and mostly considered pathological. It was not until the 1970s that they began to be seen as a single and useful family of mathematical objects that could accurately describe many physical phenomena. Mandelbrot put it as follows in the \emph{The Fractal Geometry of Nature} \cite{mandelbrot1982fractal}:

@@

Fractal geometry has been used in many fields, including physics \cite{mandelbrot1984fractals}, geology \cite{cheng1997multifractal}, biology \cite{goldberger1992fractal} and economics \cite{turiel2003multifractal}. The first connection between fractals and the natural world was in the form of tangible objects such as coast lines, trees or bronchia, but the more hidden aspects of reality were soon shown to have fractal properties as well. The path traced out by the Brownian motion of a particle, the price chart for a commodity or the level of floods in a river delta have all been studied as fractals to great effect.

@@

This problem was investigated thoroughly in the context of image compression (see for instance \cite{hart1996fractal}), but has received relatively little attention in the area of statistics, and fitting a fractal model to data rather than an image. In particular the case of data in arbitrary dimensions has been left largely uninvestigated.

\footnotetext{Mandelbrot originally defined it as a set whose topological dimension differs from its Hausdorff dimension, but retracted this definition, stating that he preferred the word to be not precisely defined.}

\subsection*{Previous research}

The most fruitful algorithms for the induction of fractal models so far have been partitioned iterated function systems and evolutionary algorithms.

Partitioned iterated function systems \cite{hart1996fractal} are a subset of iterated function systems (described in the next section) designed for fast image compression by fractal models. The compression algorithm segments an image in to two tilings of different size, and then attempts to map the larger tiles to the smaller tiles. The resulting set of contraction mappings can be iterated to approximate the original image. The approach works well for two dimensional data stored in binned form, but does not translate well to higher dimensions.

Another approach is to represent a fractal model as a point in some space (eg. a vector or a bitstring) and to search the space for a model which approximates the data by some target function using generic search algorithms. Fractal models tend to be difficult to optimize for and only specific methods show promising results. For instance, the use of evolutionary strategies combined with Hausdorff distance as a target function has been shown to be able to reconstruct fractals from data, and to fit models to common datasets.

Nevertheless, the generic nature of these search methods mean that they do not exploit any knowledge of the fractal models. We expect that the induction of fractal models can be sped up and made more accurate if we tailor the search algorithm to the model.

The investigation of iterated function systems has yielded some results that are promising, but have not resulted in an algorithm for the induction of fractal models in arbitrary dimensions. Examples include the method of moments \cite{rinaldo1994inverse}, similarity hashing \cite{hart1997similarity} and wavelet analysis \cite{struzik1996coastline}.

\section*{Methods}

In this paper we focus on numerical data in an arbitrary number of dimensions. Each instance in our dataset is represented by a fixed number of numerical features. Call our dataset $X$, our instance $x \in X$ and $x \in {\mathbb R}^d$. We will assume that our dataset represents iid samples from a probability distribution over ${\mathbb R}^d$ which we will attempt to model.

\subsection*{Iterated Function Systems}

@@

We will use the following notation. Let $S_1$, $S_2$, \ldots, $S_k$ be a set of contractive transformations on ${\mathbb R}^d$. (We will later limit the family of allowable functions, but from a mathematical point of view we only require them to be contractive). There exists a single set $Q$ that satisfies the following property:
\begin{align}
Q = S_1(Q) \cup S_2(Q) \cup \ldots \cup S_k(Q) \label{scaling law}
\end{align}

@@

\subsection*{Representing IFSs}

@@

The method described below can easily be extended to other families of transformations such as affine transformations or neural networks. The only requirement is that we must have some way of finding a transformation $F$ within this family given two ordered point sets $\{x_i\}, \{y_i\} \subseteq {\mathbb R}^d$ so that the error is minimized. For similarities there is a direct and optimal method. \cite{umeyama1991least}

\subsection*{Drawing points from an IFS distribution}

@@

It should be mentioned that the sequence of points generated this way does not represent a sequence of independent draws. Any two successive points in the sequence will satisfy the relation $x_{n+1} = S_i(x_{n})$ for some $i$. This is not true for two independently drawn points. For our experiments we generate a sequence and then shuffle the data to remove this dependency where it is undesirable.

@@

\subsection*{Code space}
\label{sec:code-space}

@@

Since each $S_i$ is contractive, we can let the length of $\sigma$ go to infinity and get 
\[
S_\sigma(X) = x_\sigma  
\] 
That is, the sequence of functions maps each $X$ to a unique point $x_\sigma$. In the case of probability measures, every probability measure is mapped to a probability measure that gives the single point $x_\sigma$ full probability mass. Figure [@!] illustrates the concept, an Iterated Function System assigns a code to each point on its support. Finite codes map to subsets of the support. Since the size of the subsets decreases exponentially we can talk about the endpoint of a finite code without losing much accuracy. We will take the point to which the origin is mapped by $S_\sigma$ as the canonical choice of the endpoint of finite $\sigma$.

Finding the endpoint for given code is straightforward, but we will also require the inverse process: given a point $x$ on or near the support of some IFS, we wish to find $\sigma$. To this end we start with the standard normal multivariate distribution and apply all maps $S_\sigma$ for all $\sigma$ of a given length $l$ giving us a set of $k^l$ normal distributions (since an affine transformation of a normal distribution is also a normal distribution). The $\sigma$ which assigns $x$ the greatest probability is chosen.

Since the variance of the distribution drops off exponentially as the length $l$ is increased, this method will fail if points are too far off the support of the IFS (even if logarithmic probabilities are used). To increase the range of the method, we fall back on the distance from the point to the mean of the distribution. If the first method cannot distinguish between codes, we choose the code whose probability distribution has its mean closest to the point.

A word of caution on codes is appropriate: the ordering is often the reverse of what our intuition suggests. If we number the three components of the Sierpinski triangle clockwise from the top 0, 1 and 2, we might number a small section of the triangle by first moving into the top subtriangle, then into the left subtriangle (2) of that and then into the right subtriangle (1) of that. Intuitively, this is the most natural way of coding sections of the Sierpinski triangle.

@@

Finally, we wish to note that a line segment, the square, the cube and higher dimensional analogues can be represented as simple IFSs of $2^d$ components. For instance, for the unit line, we map the whole onto its lower and upper half with two components. The codes of this IFS are binary strings and the subsets they code are the diadic intervals. For the square, we can map the whole onto each one of its quadrants and get quaternary codes. Using this mechanism we can always map any dataset into the bi-unit hypercube and find a straightforward coding for the instances in $2^d$-ary codes. What fractal fitting offers us is a potential for a subspace that fits the data more compactly (ie. using fewer components) and gives us a good code of lower arity.

\begin{figure}[b!]
  \centering
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{../img/sierpinski-codes.pdf}
    \caption{}
    \label{fig:sierpinski-codes}
  \end{subfigure}  
  \hspace{0.015\textwidth}
  \begin{subfigure}[b]{0.48\textwidth}
    \includegraphics[width=\textwidth]{../img/code-construction.pdf}
    \caption{}
    \label{fig:code-construction}
  \end{subfigure}

  \caption{\small (a) Length-three codes on the Sierpinski triangle and the subsets they code for. (b) The construction of a subset from its code.}
  \label{fig:codes}
\end{figure}

\subsection*{Evaluating models}

@@

Consider a dataset drawn from the Sierpinski gasket with a minute amount of two dimensional noise added. Since the Sierpinski gasket has dimension less than two, the probability that the point will still fall within the Sierpinski gasket will be zero. Thus all points will have probability zero under the Sierpinski gasket and its likelihood will be zero. Intuitively however, the Sierpinski gasket is still a good model for this data.

For this reason we will often use the Hausdorff distance to evaluate our models. The Hausdorff distance  $H(X, Y)$ between two sets $X$ and $Y$ is defined as follows:
\[ 
H'(X, Y) = \max_{x\in X} \, \min_{y\in Y} \, d(x, y) 
\]
\[
H(X, Y) = \max\left (H'(X, Y), H'(Y, X)\right )
\]
 (for infinite sets the $\max$ and $\min$ functions should be replaced with $\sup$ and $\inf$)

The most intuitive way of understanding the Hausdorff distance is as a game where the first player must get from one set to the other in as small a distance as possible, and the malevolant second player will choose his starting point to maximize that distance.

Hausdorff distance is a common tool in dealing with the fractal inverse problem \cite{nettleton1994evolutionary}. We will use it to evaluate an IFS model by generating a number of points from the model and calculating the Hausdorff distance of that set to the model. Since the Hausdorff distance has not been substantiated as a statistical measure, we can only use it as an intermediate indication of model quality. To prove the utility of IFS models we will test them on classification and dimensionality reduction tasks.

\subsection*{Dimension}

The \textit{curse of dimensionality} is a blanket term for many effects related the analysis of high dimensional data. The main cause of these problems is that as dimensionality increases, the volume of a subspace (such as the unit ball) increases exponentially with it. This means, for instance, that vast amounts of samples are required to maintain the same level of coverage in a high dimensional space as compared to a low dimensional one.

@@

\subsection*{Analogous model: Mixture of Gaussians}

To study the IFS as a statistical model, it helps to have a commonly used analogous model to compare it with. We use the Mixture of Gaussians (MoG) model. This model consists of $n$ multivariate Gaussians over ${\mathbb R}^d$ each represented by a mean $\mu \in {\mathbb R}^D$ and a $d \times d$ covariance matrix $\Sigma$. Each component also has a prior $w_i$. The complete model describes the following probability distribution:
\[
p(X) = \sum_{i \in [1,n]} w_i N(x|\mu_i, \Sigma)
\]
Since every multivariate Gaussian can also be described as an affine transformation of the standard normal distribution (mean at the origin and $\Sigma$ the identity matrix), we can also represent the mixture of Gaussians as $n$ affine transformations and $n$ weights, just as the IFS:
\[
p(X) = \sum_{i \in [1,n]} w_i S_i(N(\mu_i, \Sigma))(X)
\]
This gives us a classical model with a similar representation. Note that the representation would be equal if we used similarities for the MoG model. However since an MVN transformed with uniform scaling is rotationally symmetric that would make all the rotational parameters useless. For this reason we will allow the MoG model full affine transformations as parameters and the IFS model only similarities. Even with this discrepancy, we still consider the MoG to be a reasonable analog.

@@

\begin{description}
  \item[Expectation] Assign each data point a probability distribution over the components of the MoG model proportional to the likelihood of the model given the datapoint. In other words, each component takes a certain amount of `responsibility' for the data point.
  \item[Maximization] Estimate the parameters for each component based on the probability weights calculated in the previous step.  
\end{description}

@@

\section*{Fitting IFS models to data}

@@

\subsection*{The basic algorithm}

@@

\subsection*{Improvements}

\subsubsection*{Centering and scaling the data}

@@

\subsubsection*{Using all depths}

@@

\subsubsection*{Multiple matches}

@@

\subsubsection*{Weights}

@@

\subsubsection*{Edge cases}

@@

\section*{Results}

@@

\subsection*{Datasets}

@@

\subsection*{Model fitting and dimension scaling}

@@

\subsection*{Classification}

We explore two methods for using IFS models for classification. In the first, we train a single model for each class and combine in a straightforward manner using Bayes' rule. In the second, we learn a single model for the whole set of points, and use the codes generated by the model to perform classification.

\subsubsection*{Bayesian}

Given some data set $X = \{x_i\} \subseteq {\mathbb R}^d$ with associated classes $c_i \in \{1, 2, \ldots, m\}$, we split the data by class: $X_c = \{ x \in X \mid c_i = c\}$. We train a probability model $p_c$ for each $X_c$ and classify each new point with
\[
C(x) = \argmax_{c} p_c(x) p(c)
\]
The class prior $p(c)$ is estimated as $\frac{|X_c|}{|X|}$.

\subsubsection*{Single model}

An alternative way to construct a classifier with IFS modelling is to learn a single model for the whole dataset, ignoring class labels. For some datasets, the set of all points may present a more coherent and simple distribution than the sets split by class label. Another advantage of this approach is that we only have to learn one model, which is cheaper both in training and in evaluation.

The model provides us with a mapping from the data space (${\mathbb R}^d$) to code space ($\{1, \ldots, k\}^d$). To turn this model into a classifier we proceed as follows. We build a code tree such that each possible code represents a path from the root to the leaf node. We then traverse this tree for each point in the training set, maintaining a frequency count for each class observed at each node. Thus, the root node gives us the frequency distribution over the classes for the whole dataset and each node below it refines that to a distribution over a section of the dataset whose code starts with a given prefix. This turns each node into a very basic classifier, which simply returns the class with the highest frequency count. The root node will give us the classifier that uses nothing more than class frequencies and each node below it will use more information about the neighbourhood of the point.

@@

\section*{Conclusions}
